---
title: "Generation of Figures 1 to S2 for the DL Nuclear Segmentation Manuscript"
date: "Jan 19 2021"
authors: "Gianluca Pegoraro/George Zaki, NCI/NIH"
output: github_document
---

### Analysis initialization
Load the required libraries
```{r import-libraries}
library(tidyverse)
library(readr)
library(curl)
library(ggthemes)
```

Set `knitr` and `ggplot2` options.
```{r set-knitr-options}
knitr::opts_chunk$set(
    cache = FALSE,
    fig.path = "Output/",
    dev = c("png", "pdf"),
    message = FALSE,
    warning = FALSE
    )

theme_set(theme_bw())

theme_update(
  axis.text.x = element_text(
    angle = -45,
    hjust = 0,
    vjust = 0.5
  ),
  axis.text.y = element_text(hjust = 0)
)
```

### Read and process data and metadata
Define a function to calculate the F1 score based on the number of true positives, false positives and false negatives.
```{r calculate-F1}
calculate_F1 <- function(tp, fn, fp){
  tp/(tp + (fn + fp)/2)
}
```

Download unzip the input results of the inference experiments from Figshare if they have not been already downoladed.
```{r downlaod-data}
if(!dir.exists("Data")) {
  URL <- "https://ndownloader.figshare.com/files/24856241"
  curl_download(URL, "data.zip")
  unzip("data.zip")
}
```

Recursively read ands process all the data from all the models runs and the manual vs. automated run. This is the data for both training and testing results. 
```{r read-data, message=FALSE}
raw_tbl <- dir(path = "Data", 
               pattern = ".*maps\\.csv$", 
               recursive = TRUE,
               full.names = TRUE) %>%
  set_names() %>%
  map_df(read_csv, .id = "filename") %>%
  select(filename, 
         thres = `IoU(Threshold)`,
         TP = `Sum(all_tp)`,
         FP = `Sum(all_fp)`,
         FN = `Sum(all_fn)`,
         mAP = AP) %>%
  mutate(F1 = calculate_F1(tp = TP,
                           fn = FN,
                           fp = FP))
glimpse(raw_tbl)
```

Recode the values of the `cell_line` variable only for the model runs to make them compatible with the actual cell names. 
```{r modify-data}
levs1 <- c("MCF10A", "U2OS", "HCT116", "Eosinophils")
levs2 <-
  c("Original",
    "Technical",
    "Biological",
    "Technical_manual",
    "Biological_manual")

runs_tbl <- raw_tbl %>%
  filter(!str_detect(filename, "manual-vs-semi-auto")) %>% #only model runs
  mutate(
    filename = str_replace(filename, "BABE", "MCF10A"),
    filename = str_replace(filename, "HiTIF_Colorectal", "HCT116"),
    filename = str_replace(filename, "HiTIF_Laurent", "U2OS"),
    filename = str_replace(filename, "Manasi", "Eosinophils"),
    run = as.numeric(str_match(filename, pattern = "run([0-9]{3})")[, 2]),
    cell_line = str_match(filename, pattern = "Data/.*?/.*?/.*?/(.*?)_.*?/")[, 2],
    replicate = str_match(filename, pattern = "Data/.*?/.*?/.*?/.*?_(.*?)/")[, 2],
    cell_line = factor(cell_line, levels = levs1),
    replicate = factor(replicate, levels = levs2)
  )

glimpse(runs_tbl)
```

Recode the values of the `cell_line` variable only for the manual vs. auomated run to make them compatible with the actual cell names.
```{r modify-data-vs-manual}
manual_vs_auto_tbl <- raw_tbl %>%
  filter(str_detect(filename, "manual-vs-semi-auto")) %>% #only manual vs. semi-automated run
  mutate(
    filename = str_replace(filename, "BABE", "MCF10A"),
    filename = str_replace(filename, "HiTIF_Colorectal", "HCT116"),
    filename = str_replace(filename, "HiTIF_Laurent", "U2OS"),
    filename = str_replace(filename, "Manasi", "Eosinophils"),
    cell_line = str_match(filename, pattern = "Data/.*?/(.*?)_.*?/")[, 2],
    replicate = str_match(filename, pattern = "Data/.*?/.*?_(.*?)/")[, 2],
    cell_line = factor(cell_line, levels = levs1),
    replicate = factor(replicate, levels = levs2)
  )

glimpse(manual_vs_auto_tbl)
```

Read the run metadata table.
```{r read-metadata}
md_tbl <- read_csv("Metadata/Metadata.csv")

glimpse(md_tbl)
```

Tidy up the metadata table for visual exploration. 
```{r results='asis'}
md_explore <-
  md_tbl %>% select(model, initialization, train_set, train_size, epochs, aug, run) %>%
  arrange(model, initialization, train_set, train_size, epochs, aug, run)

knitr::kable(md_explore)
```

Reorder some factors and join the model runs table with the metadata table. Keep only the data for the test dataset runs and not for the training/evaluation dataset runs. 
```{r join-metadata}
levs3 <- c("Random", "Pre-initialized")
levs4 <-
  c("MCF10A",
    "Eosinophils",
    "MCF10A_HCT116",
    "MCF10A_HCT116_U2OS",
    "All",
    "BBBC038",
    "CellPose")
levs5 <- c("MRCNN", "FPN2-WS", "Jacobkie", "CellPose", "Kaggle_5th")


runs_tbl <- runs_tbl %>%
  left_join(md_tbl, by = c("run")) %>%
  mutate(
    initialization = factor(initialization, levels = levs3),
    train_set = factor(train_set, levels = levs4),
    train_size = factor(train_size),
    run = factor(run),
    aug = factor(aug),
    epochs = factor(epochs),
    model = factor(model, levels = levs5)
  ) 

glimpse(runs_tbl)
```

Separate testing datasets from training datasets. 
```{r filter-testing}
testing_tbl <- runs_tbl %>%
  #only test datasets
  filter(replicate != "Original") %>% 
  # only the `biological` replicate for MCF10 cells
  filter(!(cell_line == "MCF10A" & replicate == "Technical")) 

glimpse(testing_tbl)
```

Select only the F1 values for IoU = 0.7 (As in [the KAGGLE '18 nucleus challenge paper](https://paperpile.com/shared/xZiVUo)).
```{r filter-testing-zero-seven}
testing_07_tbl <- testing_tbl %>%
             filter(round(thres, digits = 2) == 0.70, 
                    replicate %in% c("Technical", "Biological"))
```

### Plotting and Table Generating functions
```{r gen-line}
gen_line <- function(df) {
  df %>%
  ggplot(aes(
    x = thres,
    y = F1,
    color = cell_line
  )) +
    geom_line() +
    geom_vline(xintercept = 0.7,
               linetype = "longdash",
               color = "grey60") +
    facet_wrap(vars(model)) +
    scale_linetype_discrete("Replicate") +
    scale_color_few(name = "Cell Line") +
    xlab("IoU Threshold") +
    ylab("F1 Score")
}


gen_point <- function(df, varx, x_legend){
  df %>% 
    ggplot(aes(x = {{varx}},
               y = F1,
               color = model)) +
    geom_point(alpha = 0.7,
               size = 3) +
    scale_y_continuous(lim = c(0, 1.05), 
                       breaks = seq(0, 1, 0.1)) +
    scale_color_few(name = "Model") +
    facet_wrap(vars(.data$cell_line),
                    nrow = 1) +
    xlab(x_legend) +
    ylab("F1 Score @ IoU 0.7")
}
```

```{r gen-table}
gen_table <- function(df, varx){
  df %>%
    filter(round(.data$thres, digits = 2) %in% c(0.70, 0.9)) %>%
    select(.data$model, .data$cell_line, {{varx}}, .data$F1) %>%
    pivot_wider(
      names_from = c(.data$model, {{varx}}),
      values_from = .data$F1,
      names_sep = "-"
    )
}
```

### Figure 2
Create a subset of testing F1 results for the baseline models. 
```{r filter-2B}
fig_2B <- testing_tbl %>%
  filter(run %in% c(3, 4),
         replicate %in% c("Technical", "Biological"))
```

Plot the baseline model training performance using only MCF10A cells images for training and using random weights (Fig. 2B).
```{r Fig2B, echo = F}
gen_line(fig_2B)
```

Generate and save a table for the F1 results shown in Fig 2B at the 0.7 and 0.9 thresholds (Table1).
```{r table1}
table1 <- gen_table(fig_2B, thres)

write_csv(table1, path = "Output/Table1.csv")
```

### Figure 3
Create a subset of F1 testing results for the experiments where we tested changes to the model inizialization.
```{r filter-3A}
fig_3A <- testing_07_tbl %>% 
  filter(run %in% c(1, 3, 4, 5))
```

Plot only the effect of changes to initialization (Fig. 3A).
```{r Fig3A, echo=F}
gen_point(fig_3A, initialization, "Initialization")
```

Generate and save a table for the F1 results shown in Fig 3A at the 0.7 threshold (Table2).
```{r table2}
table2 <- gen_table(fig_3A, initialization)

write_csv(table2, path = "Output/Table2.csv")
```

Create a subset of F1 testing results for the experiments where we tested changes to the composition of the training datasets (Fig.3B).
```{r filter-3B}
fig_3B <- testing_07_tbl %>% 
  filter(run %in% c(1, 5, 6, 10, 29, 30, 65, 66)) 
```

Plot only the effect of incremental training (Fig. 3B).
```{r Fig3B, echo = F}
gen_point(fig_3B, train_set, "Training Set")
```

Generate and save a table for the F1 results shown in Fig 3B at the 0.7 threshold (Table3).
```{r table3}
table3 <- gen_table(fig_3B, train_set)

write_csv(table3, path = "Output/Table3.csv")
```

### Figure 4
Create a subset of F1 testing results for the experiments where we tested changes in the length of model training (Fig.4A).
```{r filter-4A}
fig_4A <- testing_07_tbl %>% 
  filter(run %in% c(6, 10, 12, 16, 17, 18, 67, 68))
```

Plot only the effect of incremental length of training (Fig. 4A).
```{r Fig4A, echo=F}
gen_point(fig_4A, epochs, "Number of Training Cycles")
```

Generate and save a table for the F1 results shown in Fig 4A at the 0.7 threshold (Table4).
```{r table4}
table4 <- gen_table(fig_4A, epochs)

write_csv(table4, path = "Output/Table4.csv")
```

Create a subset of F1 testing results for the experiments where we tested changing the number of augmented ROI patches used for training (Fig.4B).
```{r filter-4B}
fig_4B <- testing_07_tbl %>% 
  filter(run %in% c(6, 10, 11, 19, 20, 45:49))
```

Plot only the effect of incremental size of the training set (Fig. 4B).
```{r Fig4B, echo=F}
gen_point(fig_4B, train_size, "Number of Augmented ROIs Used for Training ")
```

Generate and save a table for the F1 results shown in Fig 4B at the 0.7 threshold (Table5).
```{r table5}
table5 <- gen_table(fig_4B, train_size)

write_csv(table5, path = "Output/Table5.csv")
```

Create a subset of F1 testing results for the experiments where we tested changing the augmentation strategy used for image preprocessing (Fig.4C).
```{r filter-4C} 
fig_4C <- testing_07_tbl %>% 
  filter(run %in% c(6, 10, 22:25, 27, 62, 69:72))
```

Plot only the effect of changing the augmentation strategies (Fig. 4C).
```{r Fig4C, echo=F}
gen_point(fig_4C, aug, "Augmentation")
```
Generate and save a table for the F1 results shown in Fig 4C at the 0.7 threshold (Table6).
```{r table6}
table6 <- gen_table(fig_4C, aug)

write_csv(table6, path = "Output/Table6.csv")
```

Create a subset of F1 testing results for the experiments where we tested changing the number of augmented ROI patches used for training augmentations experiments and inizializations several times (i.e. repeats) (Fig.4D).
```{r filter-4D}
fig_4D <- testing_07_tbl %>% 
  filter(run %in% c(6, 8, 10, 26, 32, 33:43, 50:64, 73:83))
```

Plot the effect of changing the augmentation strategies and inizializations several times (Fig. 4D).
```{r Fig4D, echo = F}
gen_point(fig_4D, 
          interaction(aug, initialization, sep = "/"), 
          "Augmentation/Initialization") +
    stat_summary(
    fun.data = "mean_cl_normal",
    geom = "crossbar",
    width = 0.5
  )
```

Generate and save a table for the F1 results mean and 95% Confidence Interval as shown in Fig 4C at the 0.7 threshold (Table7).
```{r table7}
table7 <- fig_4D %>%
  group_by(cell_line, aug, initialization, model) %>%
  summarize(CI_95 = list(Hmisc::smean.cl.normal(F1))) %>%
  unnest_wider(CI_95) %>%
  rename(F1_mean = Mean,
         CI_95_lower = Lower,
         CI_95_upper = Upper)

write_csv(table7, "Output/Table7.csv")
```

### Figure 5
Create a subset of F1 testing results for the final models (Runs 10 for FPN2-WS and 37 for MRCNN) and comparison with the out-of-the-box Jacobkie model (Run 44) (Fig. 5B).
```{r filter-5B}
fig_5B <- testing_tbl %>% 
  filter(run %in% c(10, 37, 44),
         replicate %in% c("Technical", "Biological"))
```

Plot the performance for the final MRCNN and FPN2-WS models comparing it with Jacobkie. This is Fig. 5B.
```{r Fig5B, echo = F}
gen_line(fig_5B)
```

Generate and save a table for the F1 results shown in Fig 5B at the 0.7 and 0.9 thresholds (Table8).
```{r table8}
table8 <- gen_table(fig_5B, thres)

write_csv(table8, path = "Output/Table8.csv")
```

### Figure S1B
Create a subset of F1 testing results for the final models (Runs 10 for FPN2-WS and 37 for MRCNN) and comparison with the out-of-the-box Jacobkie model (Run 44) using both semi-automated and manual labelling (Fig. S1B).
```{r filter-S1B}
fig_S1B <- testing_tbl %>% 
               filter(run %in% c(10, 37, 44)) %>%
               mutate(annotation = ifelse(str_detect(replicate, "manual"), "Manual", "Semi-Automated"))
  
```

Plot the performance for the final MRCNN and FPN2-WS models comparing it with Jacobkie (Run 44) using both semi-automated generated labels and manually annotated labels. This is Fig. S1B.
```{r FigS1B, echo = F}
fig_S1B %>% 
  ggplot(aes(x = thres, 
             y = F1,
             color = annotation)) +
  geom_line() +
  geom_vline(xintercept = 0.7, 
             linetype = "longdash",
             color = "grey60") +
  facet_grid(vars(cell_line), vars(model)) +
  scale_color_few(name= "Annotation") +
  xlab("IoU Threshold") +
  ylab("F1 Score")
```

Generate and save a table for the F1 results shown in Fig S1B at the 0.7 and 0.9 thresholds.
```{r tableS3}
tableS3 <- fig_S1B %>%
  filter(round(thres, digits = 2) %in% c(0.70, 0.9)) %>%
  select(model,annotation, cell_line, thres, F1) %>%
  pivot_wider(names_from = c(model, annotation, thres),
              values_from = F1,
              names_sep = "-")

write_csv(tableS3, path = "Output/TableS3.csv")
```
### Figure S2A
Plot Fig S2A: comparison of the manual vs. semi-automated labelling strategy. 
```{r FigS2A, echo=FALSE}
manual_vs_auto_tbl %>%
  ggplot(aes(x = thres,
             y = F1)) +
  geom_line() +
  geom_vline(xintercept = 0.7, 
             linetype = "longdash",
             color = "grey60") +
  facet_wrap(vars(cell_line),
             nrow = 1) +
  xlab("IoU Threshold") +
  ylab("F1 Score")
```

### Figure S2B
Create a subset of F1 testing results for the final models (Runs 10 for FPN2-WS and 37 for MRCNN)  using either the F1 training or testing inference results (Fig. S2B).
```{r filter-FigS2B}
fig_S2B <- runs_tbl %>%
  filter(run %in% c(10, 37),
         replicate %in% c("Original", "Technical", "Biological")) %>%
  filter(!(cell_line == "MCF10A" & replicate == "Technical")) %>%      
  mutate(replicate = if_else(str_detect(replicate, "Original"), "Training", "Testing")) %>%
  mutate(replicate = factor(replicate, levels = c("Training", "Testing"))) 
```

Plot Fig. S2B.
```{r FigS2B, echo=FALSE}
fig_S2B %>%
  ggplot(aes(x = thres,
             y = F1,
             color = replicate)) +
  geom_line() +
    geom_vline(xintercept = 0.7, 
             linetype = "longdash",
             color = "grey60") +
  facet_grid(vars(cell_line),
             vars(model)) +
  scale_color_few(name= "Dataset") +
  xlab("IoU Threshold") +
  ylab("F1 Score")
```

Generate and save a table for the F1 results shown in Fig S2B at the 0.7 and 0.9 thresholds (Table S4).
```{r tableS4}
tableS4 <- fig_S2B %>%
  filter(round(thres, digits = 2) %in% c(0.70, 0.9)) %>%
  select(model, replicate, cell_line, thres, F1) %>%
  pivot_wider(names_from = c(model, replicate, thres),
              values_from = F1,
              names_sep = "-")

write_csv(tableS4, path = "Output/TableS4.csv")
```

Bonus: comparison of MRCNN and FPN-WS2 performance with (Cellpose) [https://cellpose.readthedocs.io/en/latest/]
```{r filter-FigCellPose}
cell_pose <- testing_tbl %>% 
  filter(run %in% c(10, 37, 44, 84),
         replicate %in% c("Technical", "Biological"))
```

```{r FigCellPose, echo=FALSE}
gen_line(cell_pose)
```
```{r, tableCellPose}
cell_pose_table <- gen_table(cell_pose, thres)

write_csv(cell_pose_table, path = "Output/cell_pose_table.csv")
```

```{r sessionInfo, results='markup'}
sessionInfo()
```